Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-05-16 ms.tex
forgot some of davids edits

Our estimated timing for biopoesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology . Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be mostly compatible with our estimating timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga.

@@

An additional related caveat is that our calculations assume that an impact that vaporizes the ocean also sterilizes the planet, but this may not be the case because microbes could have possibly survived in the deep subsurface \citep{Sleep_1989,Grimm_2018}. If ocean-vaporization does not destroy the biosphere, then our results would favor an origin of life earlier than we predict in Figure \ref{fig:time_and_mass}, and a life-starting impact would also be more probable than we have estimated (Figure \ref{fig:probabilities_of_impacts}).

@@

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life in an RNA-first processes. We make use of results of \citet{Wogan_2023}, who find that significant origin of life molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by a smaller ocean-vaporizing collision that sterilizes the planet.

@@

If life began in an RNA-first process after an impact, then this work suggests that origin of life on Earth was a fairly likely event (at least a 20\% chance) and was indeed strongly favored (92\% chance) under optimistic assumptions. Given that rocky planets form from accretionary impacts, our work supports and optimistic outlook for life on exoplanets orbiting sun-like stars and future searches for biosignatures.

>>>>

Our estimated timing for biopoesis is compatible with the earliest well-accepted geologic evidence of life in the form of stromatolites in the 3.5 Ga Pilbara block of western Australia \citep{Walter_1980,Buick_1981,Van_2018}. Older geologic evidence of life exists, such as a $> 3.7$ Ga black shale metamorphosed to graphite with negative $\delta^{13}$C characteristic of biology \citep{Rosing_1999,Ohtomo_2014}. Also, \citet{Bell_2015} discovered graphite inclusions in a 4.1 Ga zircon grain with $^{13}$C depletions compatible with biological activity, but this is a tentative biosignature because there are abiotic mechanisms for making isotopically light carbon \citep{Javaux_2019}. If the \citet{Bell_2015} finding is indeed a sign of life, then it would be generally compatible with our estimating timing in Figure \ref{fig:time_and_mass}, because we find that there is only a $\sim 10\%$ chance of life starting after 4.1 Ga.

@@

An additional related caveat is that our calculations assume that an impact that vaporizes the ocean also sterilizes the planet, but this may not be the case because microbes might have survived in the deep subsurface \citep{Sleep_1989,Grimm_2018}. If ocean-vaporization does not destroy the biosphere, then our results would favor an origin of life earlier than we predict in Figure \ref{fig:time_and_mass}, and a life-starting impact would also be more probable than we have estimated (Figure \ref{fig:probabilities_of_impacts}).

@@

We use Monte-Carlo simulations of Earth's impact history to determine the most likely timing for the origin of life that requires early RNA synthesis. We use the results of \citet{Wogan_2023}, who find that significant origin of life precursor molecules, such as HCN, are produced in the Hadean atmosphere after a $> 4 \times 10^{20}$ kg impact in an optimistic case, and after a $> 5 \times 10^{21}$ kg impact in a pessimistic scenario. We consider both possibilities, and assume such impacts can cause life's origin so long as they are not followed by a smaller ocean-vaporizing collision that sterilizes the planet.

@@

If life began after an impact that generated a reducing atmosphere, then this work suggests that origin of life on Earth was fairly likely (at least a 20\% chance) and was indeed strongly favored (92\% chance) under optimistic assumptions. Given that rocky planets form from accretionary impacts, our work supports and optimistic outlook for life on exoplanets orbiting sun-like stars and future searches for biosignatures.
